---
title       : Cancer cases in Switzerland from 1984 to 2008
subtitle    : Incidence cases and annual rate per 100 000
author      : Lukas 
framework   : deckjs        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [bootstrap]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

# Cancer in Switzerland from 1984 to 2008


### Course project of the coursera class "[Developing Data Products](https://class.coursera.org/devdataprod-006)"

Incidence number and average annual rate per 100 000 persons

Visualized with [Shiny](http://shiny.rstudio.com/): https://ijon.shinyapps.io/shiny/

---

## Displays development of cancer frequencies from 1984 to 2008

* displays either
  * cancer incidence **number**:  
    absolute number within given period (e.g. 1984 to 1988)
  * cancer indicence **rate**:  
    average annual number of new cancer cases per 100 000 persons in a given period
* select type of cancer, for example: 
  * Liver
  * Eye
  * or total number, summarized over all types of cancer
* select sex: display results
  * only for women
  * only for men
  * or for both

Results are displayed as a stacked bar graph:  
one bar for each 5-year period between 1984 and 2008.

--- .class #id

## Data source: Cancer epidemiology of the Swiss Federal Statistical Office

* [Cancer epidemiology of the Swiss Federal Statistical Office](http://www.bfs.admin.ch/bfs/portal/en/index/infothek/erhebungen__quellen/blank/blank/kbs/01.html)
* Get the data [here](http://www.pxweb.bfs.admin.ch/dialog/varval.asp?ma=px-d-14-2B11&ti=Krebs%3A+Anzahl+und+j%E4hrliche+Rate+der+Neuerkrankungen+und+Todesf%E4lle+nach+Altersgruppe%2C+Geschlecht%2C+Region+und+Krebslokalisation%2C+1984%2D2008&path=../Database/German_14%20-%20Gesundheit/14.2%20-%20Gesundheit%20der%20Bev%F6lkerung/&search=KREBS&lang=1)


--- .class #id 

## Server side calculations: 
   
Preparation: **load, clean, calculate percentage**   
For each page reload: **select rows, and display with ggplot2**

### Data preparation

* load the data from a csv file
* remove rows with no data
* change empty cells of sex-specific cancer types (e.g. Testicle cancer for women) from NA to 0
* Calculate percentage of men and women for a given cancer type using package ```ddplyr```.

```{r echo=FALSE}
dataset <- read.csv( 'assets/data/cancer.csv', header=TRUE, sep=",", na.strings=c("", "*"))

dataset <- dataset[ !is.na(dataset$Period), ]
dataset[ is.na(dataset$Incidence.num), "Incidence.num" ] <- 0
dataset[ is.na(dataset$Incidence.rate), "Incidence.rate" ] <- 0

# Replace label for 'Cancer.total' with 'All cancers':
levels(dataset$Cancer)[levels(dataset$Cancer)=='Cancer.total'] <- 'All cancers'

# Calculate percentage of Incidates for men/woman

library(plyr)

calc.percent <<- function(num) {
  sapply( num, function(x) { ifelse( is.na(x), NA, paste(round( 100 * x ), '%', sep='') ) })
}

dataset <- ddply( dataset, .(Cancer,Period), summarise, 
                  Sex=Sex, Incidence.num = Incidence.num, 
                  Inc.num.perc =   calc.percent( Incidence.num / sum(Incidence.num, na.rm=T) ),
                  Incidence.rate = Incidence.rate, 
                  Inc.rate.perc =  calc.percent( Incidence.rate / sum(Incidence.rate, na.rm=T) )
)

# Prepare a function that selects the correct rows from the dataset. 
# To be used in the shinyServer function.
select.data <- function(input) {
  if (input$sex == 'Only women') {
    selected.data <- dataset[dataset$Cancer==input$cancer & dataset$Sex=='Women', ]
  } else if (input$sex == 'Only men') {
    selected.data <- dataset[dataset$Cancer==input$cancer & dataset$Sex=='Men', ]
  } else {
    selected.data <- dataset[dataset$Cancer==input$cancer, ]
  }
  selected.data
}

colors <- c("cyan", "orange")      

library(ggplot2)
```

### Calculation during each page reload

* select correct rows (depending on cancer type, sex)
* display correct variables: either incidence *number* or incidence *rate*
* create a ggplot2 plot with correct axis, titles (see next slide)

--- .class #id 

## Result: ggplot diagram

Say the user chose: *Display incidence of all types of cancers for both men and women*:

```{r fig.width=12}
input <- list(sex="both", cancer="All cancers")  # simulate Shiny input parameter
ggplot(data=select.data(input), aes(x=Period, y=Incidence.num, fill=Sex, label=Inc.num.perc, ymax=40000)) + 
  scale_fill_manual(values=colors) +  ylab("Incidence") +
  theme(axis.title=element_text(size=20), axis.text=element_text(size=16), legend.title=element_text(size=20),
    legend.text=element_text(size=16)) +
  guides(fill=guide_legend(title=NULL)) + geom_bar(stat="identity") + geom_text(position="stack", vjust=2)
```
